\documentclass[a4paper,12pt,oneside,openright]{article}
\usepackage[pdftex] {graphicx}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{subfig}
\usepackage{chapterbib}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{wrapfig}

\addtolength{\hoffset}{-1cm}
\addtolength{\textwidth}{2.5cm}

\lstdefinelanguage{J}[]{C}
{morekeywords={},
sensitive=false}
\lstset{language=C}
% shadowbox come frame
% rulesepcolor=\color{Gray}
\lstset{tabsize=3, breaklines, commentstyle=\texttt}
\lstset{showstringspaces=false}
\lstset{basicstyle=\small}
\lstset{emph={},emphstyle=\textit}

\title{Performance evaluation of parallel sorting algorithms for large data sets}
\author{Nicola Desogus, Paolo Giangrandi, Fabio Luporini}

\begin{document}
\maketitle
\tableofcontents
\pagebreak

\include{state_art}

\input{algorithms}

\section{Structuring the project: a top-down approach}
\subsection{Abstraction Layers}
\input{framework}
\subsection{Sorting Algorithms}
\input{terminology}
\input{mergesort}
\input{kmerge}
\input{lbmergesort}
\input{lbkmergesort}
\input{quicksort}
\input{bitonicsort}
\input{bucketsort}
\input{samplesort}
\subsection{Sorting Framework}
\input{sorting_framework}
\subsection{DAL: Data Abstraction Layer}
\input{data}

\input{test_env}

\input{futureworks}

\pagebreak

\begin{thebibliography}{13}
\bibitem{FERR}{Paolo Ferragina. \textit{The magic of Algorithms!.} Lectures notes in Algorithm Engineering, 2010}
\bibitem{CPSA}{Nancy M. Amato	 and Ravishankar Iyer and Sharad Sundaresan and Yan Wu.\textit{A Comparison of Parallel Sorting Algorithms on Different Architectures.} Technical Report 98-029, Department of Computer Science, Texas A$\&$M University, January 1996. }
\bibitem{CSPA2}{Marc Moreno Maza. \textit{CS855: Parallel Sorting Algorithms.} Technical Report, Department of Computer Science, The University of Western Ontario, February 2008.}
\bibitem{NPSA}{David R. Cheng and Alan Edelman and John R. Gilbert and Viral Shah. \textit{A Novel Parallel Sorting Algorithm for Contemporary Architectures.} In ALENEX06, May 2007.}
\bibitem{VANN}{Marco Vanneschi. \textit{Architettura degli Elaboratori.} Plus-Pisa University Press, 2009.}
\bibitem{MPI}{http://www.mpi-forum.org/docs/mpi-11-html/node41.html$\#$Node41}
\bibitem{SAN}{Kurt Mehlhorn and Peter Sanders. \textit{Algorithms and Data Structures: The Basic Toolbox.} Springer-Verlag, June 2008.}
\bibitem{MVAPICH2-MAPPING}{http://mvapich.cse.ohio-state.edu/support/user$\_$guide$\_$mvapich2-1.5.1.html$\#$x1-330006.2}
\end{thebibliography}

\pagebreak

\appendix
\include{appendix}
\include{appendixB}

\end{document}
